---
  output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.align = "center",
  dpi = 300
)
```

## Introduction

The **ibdsegments** package offers functions related to identity by descent (IBD) probability for pedigree members. Both the discrete case (identity coefficients) and the continuous case (segments length) are treated using a Hidden Markov Model (HMM) approach. Key functionality includes:

-   Computing identity coefficients for two pedigree members
-   Random sampling of IBD segments
-   Obtaining the full probability distribution of the fraction (or total length) of an autosome that is IBD between two or more pedigree members
-   Convolution of IBD distributions using the FFT

## Getting started
```{r}
library(ibdsegments)
```

#### Identity coefficients
The `d_ibd` function may be used to compute identity coefficients. The example shows how to compute the `kappa` coefficients for half siblings.

```{r}
ped_hs <- pedtools::halfSibPed()

d_ibd(0, pedigree = ped_hs, coefficients = "kappa")
d_ibd(1, pedigree = ped_hs, coefficients = "kappa")
```

Besides `kappa`, other identity coefficients supported include:

* `ibd`: 0, 1, or 2 whenever all selected pedigree members jointly share this number of founder allele labels
* `identity`: Jacquard's 9 condensed identity coefficients 
* `detailed`: the 15 detailed identity coefficients

For example, the inbreeding coefficient may be computed as an `ibd` state for a single pedigree member.
```{r}
ped_inbred <- pedtools::fullSibMating(n = 1)
d_ibd(ibd = 1, pedigree = ped_inbred, ids = 5, coefficients = "ibd")
```
The probability that three siblings are jointly double ibd is also easily computed using the `d_ibd` function.
```{r}
ped_3fs <- pedtools::nuclearPed(nch = 3)
d_ibd(ibd = 2, pedigree = ped_3fs, coefficients = "ibd")
```

#### Random sampling of IBD segments

The identity coefficients computed above are IBD probabilities at single positions on a chromosome. Taking a continuous view, the fraction of the chromosome that is in each of the states is in expectation equal to these IBD probabilities. The `r_cibd` functions implements random sampling of continuous IBD:

```{r}
set.seed(1)
r_cibd(n = 1, pedigree = ped_hs, coefficients = "kappa", chromosome_length = 100)
```

### IBD distribution
The `total_ibd_dist` function obtains the full distribution of the total length of IBD segments across a chromosome. For example, we may obtain this distribution for half siblings on a chromosome with a length of 100 cM.
```{r}
d_hs <- total_ibd_dist(ped_hs, chromosome_length = 100)
d_hs
```
The distribution has two point masses (no IBD at all and fully IBD) and admits a density function otherwise. A plot includes both components.
```{r}
plot(d_hs)
```
Utility functions for computing the expectation, variance and standard deviation of the distributions are also available. These functions use numerical integration.
```{r}
E(d_hs)
sd(d_hs)
```

## Convolution of IBD distributions

The convolution of the total IBD distribution across chromosomes is obtained when the `chromosome_length` parameter has length greater than 1.

```{r}
d_hs_conv <- total_ibd_dist(ped_hs, 
                       chromosome_length = c(250, 200, 150, 150, 100))
plot(d_hs_conv)
```

Because the number of point masses may increase quickly, by default any point mass below `1e-9` is removed.

```{r}
L <- c(267.77, 251.73, 218.31, 202.89, 197.08, 186.02, 178.4, 161.54, 
       157.35, 169.28, 154.5, 165.49, 127.23, 116, 117.32, 126.59, 129.53, 
       116.52, 106.35, 107.76, 62.88, 70.84)

d_hs_full_conv <- total_ibd_dist(ped_hs, chromosome_length = L)
d_hs_full_conv
plot(d_hs_full_conv)
```
